Methods for Large Scale Learning to Rank

### 2014-11-21 report/Chapters/Conclusions.tex
laatste werk

Moreover, we found that the addition of a normalisation preprocessing step to the ListNet procedure can greatly improve the ranking accuracy of the ListNet training procedure after the first five iterations. This suggests that the addition of a normalisation preprocessing step can reduce the number of training iterations needed for convergence. Lin stated in his essay \cite{Lin2013} that MapReduce is often good enough for tasks that are not-amenable to the MapReduce model. Lin motivates this statement in the context of iterative algorithms with the observation that these iterative algorithms can often be optimised in such a way that less iterations are needed for convergence. Our preprocessing step can improve the convergence rate of the ListNet training iteration, and therefore fits into Lin's point of view.\\

Most importantly, we found the training time of our cluster version of ListNet to grow better than linearly in terms of data size increase. This shows that the cluster implementation of ListNet can be used to scale the ListNet training procedure to arbitrarily large data sets, given that enough data nodes are available for computation.\\

>>>>

Moreover, we found that the addition of a normalisation preprocessing step to a data set that is not yet normalised can greatly improve the convergence of the ranking accuracy of the ListNet training procedure. Lin stated in his essay \cite{Lin2013} that MapReduce is often good enough for tasks that are not-amenable to the MapReduce model. Lin motivates this statement in the context of iterative algorithms with the observation that these iterative algorithms can often be optimised in such a way that less iterations are needed for convergence. Our preprocessing step can improve the convergence rate of the ListNet training iteration, and therefore fits into Lin's point of view.\\

Most importantly, we found the training time of our cluster version of ListNet to grow linearly in terms of data size increase. This shows that the cluster implementation of ListNet can be used to scale the ListNet training procedure to arbitrarily large data sets, given that enough data nodes are available for computation.\\

### 2014-11-21 report/Chapters/Results.tex
laatste werk

Table \ref{tbl:accuracy_comparison} shows the mean and the standard deviation of the ranking accuracy in \ac{NDCG}@10 on the first data fold for our own cluster implementation (with and without the preprocessing step described in section \ref{ssec:preprocessing}) of ListNet, the RankLib implementation of ListNet, as well as the official LETOR evaluation results of ListNet as reported on the LETOR 3.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} websites. Note that the official LETOR evaluation results are structurally higher compared to the results obtained with RankLib and our cluster implementation, which can be attributed to the fact that the official evaluation runs use as many ListNet iterations as needed for convergence on the validation fold, while we limited our experiment to five iterations. The official evaluation results of the LETOR benchmarks do not report the standard deviation of their experimental results and it is not mentioned in the LETOR 3.0 and 4.0 papers \cite{Qin2010, Qin2013} how many times their experiments were repeated. For the ListNet step size hyperparameter we choose a value of $0.0001$, which is the default value step size value in RankLib, for both the cluster and the RankLib runs. It is not clear what step size was used in the official evaluation runs. We expect, given that LETOR aims to compare ranking methods in terms of their ranking potential, that the official LETOR benchmark results optimises the step size hyperparameter on the validation set and reported the performance on the optimal step size found with this procedure. Using this procedure one is expected to find a higher ranking accuracy than one would expect to find using a default step size value. Hyperparameter optimisation on the validation set would however not be feasible for our cluster implementation of ListNet given the Hadoop job scheduling overhead of the cluster.\\

@@

A remarkable finding is that the cluster implementation with the normalisation preprocessing procedure shows better ranking accuracy than the RankLib version of ListNet after five iterations. It seems to be the case that the normalisation procedure enables ListNet to converge faster. It has been shown in literature, that the gradient descent optimisation procedure, which is also being used in ListNet, converges faster on normalised data \cite{Ng1999}.

@@

@@

\begin{figure}
\centering
\includegraphics[trim=5cm 5cm 0cm 5cm, scale=0.7]{gfx/speedup_faceted.pdf}
\caption{Processing time of a single ListNet training iteration as a function of the number of data nodes in a cluster}
\label{fig:speedup_train_time}
\end{figure}

\begin{figure}
\centering
\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.8]{gfx/processing_speed_single_logx.pdf}
\caption{Processing speed of a single ListNet training iteration on various data sets}
\label{fig:listnet_processing_speed}
\end{figure}

>>>>

Table \ref{tbl:accuracy_comparison} shows the mean and the standard deviation of the ranking accuracy in \ac{NDCG}@10 on the first data fold for our own cluster implementation (with and without the preprocessing step described in section \ref{ssec:preprocessing}) of ListNet, the RankLib implementation of ListNet, as well as the official LETOR evaluation results of ListNet as reported on the LETOR 3.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor3baseline.aspx} and LETOR 4.0 \footnote{http://research.microsoft.com/en-us/um/beijing/projects/letor/letor4baseline.aspx} websites. Note that the official LETOR evaluation results are structurally higher compared to the results obtained with RankLib and our cluster implementation, which can be attributed to the fact that the official evaluation runs use as many ListNet iterations as needed for convergence on the validation fold, while we limited our experiment to five iterations. The official evaluation results of the LETOR benchmarks do not report the standard deviation of their experimental results and it is not mentioned in the LETOR 3.0 and 4.0 papers \cite{Qin2010, Qin2013} how many times their experiments were repeated. For the ListNet step size hyperparameter we choose a value of $1*10^{-4}$, which is the default value step size value in RankLib, for both the cluster and the RankLib runs. It is not clear what step size was used in the official evaluation runs. We expect, given that LETOR aims to compare ranking methods in terms of their ranking potential, that the official LETOR benchmark results optimises the step size hyperparameter on the validation set and reported the performance on the optimal step size found with this procedure. Using this procedure one is expected to find a higher ranking accuracy than one would expect to find using a default step size value. Hyperparameter optimisation on the validation set would however not be feasible for our cluster implementation of ListNet given the Hadoop job scheduling overhead of the cluster.\\

@@

A remarkable finding is that the cluster implementation with the normalisation preprocessing procedure shows better ranking accuracy than the RankLib version of ListNet after five iterations. It seems to be the case that the normalisation procedure enables ListNet to converge faster. It has been shown in literature that some neural network training algorithms converge faster on normalised data \cite{Ng1999}, however, this has not been researched in a Learning to Rank setting.\\

@@

Based on our own implementation of the normalisation preprocessing step that we added to the RankLib code, we further explored the effects of normalisation on the convergence rate of ListNet. We conducted experiments on the LETOR 3.0 datasets HP2003, NP2003 and TD2003 and on the MSLR-WEB10k data set. The LETOR 3.0 datasets are in published form already normalised on query-level, meaning that the $max(x)$ and $min(x)$ terms in the normalisation formula $x^{'} = \frac{x-min(x)}{max(x)-min(x)}$ are calculated per query instead of on the whole data set. The MSLR-WEB10k data set is not published in a normalised way. Looking at Figures \ref{fig:convergence_hp2003}, \ref{fig:convergence_np2003} and \ref{fig:convergence_td2003} that represent the convergence on HP2003, NP2003 and TD2003 respectively, we see that there is no difference in convergence properties between ListNet applied on globally normalised and query-level normalised data. Figure \ref{fig:convergence_mslrweb10k} shows that the ListNet convergence increases when normalisation preprocessing is applied to the unnormalised data set MSLR-WEB10k.

\begin{figure}
\centering
\includegraphics[trim=5cm 5cm 5cm 5cm, scale=0.85]{gfx/convergence_hp2003.pdf}
\caption{Convergence of ListNet on query-level and globally normalised versions of HP2003}
\label{fig:convergence_hp2003}
\end{figure}

\begin{figure}
\centering
\includegraphics[trim=5cm 5cm 5cm 5cm, scale=0.85]{gfx/convergence_np2003.pdf}
\caption{Convergence of ListNet on query-level and globally normalised versions of NP2003}
\label{fig:convergence_np2003}
\end{figure}

\begin{figure}
\centering
\includegraphics[trim=5cm 5cm 5cm 5cm, scale=0.85]{gfx/convergence_td2003.pdf}
\caption{Convergence of ListNet on query-level and globally normalised versions of TD2003}
\label{fig:convergence_td2003}
\end{figure}

\begin{figure}
\centering
\includegraphics[trim=5cm 5cm 5cm 5cm, scale=0.85]{gfx/convergence_mslr_web10k.pdf}
\caption{Convergence of ListNet on normalised and unnormalised versions of MSLR-WEB10k}
\label{fig:convergence_mslrweb10k}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[trim=0cm 5cm 0cm 5cm, scale=0.7]{gfx/speedup_faceted.pdf}
\caption{Processing time of a single ListNet training iteration as a function of the number of data nodes in a cluster}
\label{fig:speedup_train_time}
\end{figure}

\begin{figure}
\centering
\includegraphics[trim=5cm 5cm 0cm 5cm, scale=0.8]{gfx/processing_speed_single_logx.pdf}
\caption{Processing speed of a single ListNet training iteration on various data sets}
\label{fig:listnet_processing_speed}
\end{figure}
